
---
title: 'Big Data Bowl'
output:
  html_document:
    theme: simplex
    highlight: pygments
    #css: "../css/note-style.css"
  pdf_document: default
---


First, we import all of the data.
```{r}
#getwd()

games <- read.csv("../../nfl-big-data-bowl-2024/games.csv")
players <- read.csv("../../nfl-big-data-bowl-2024/players.csv")
plays <- read.csv("../../nfl-big-data-bowl-2024/plays.csv")
tackles <- read.csv("../../nfl-big-data-bowl-2024/tackles.csv")
tracking_week_1 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_1.csv")
tracking_week_2 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_2.csv")
tracking_week_3 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_3.csv")
tracking_week_4 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_4.csv")
tracking_week_5 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_5.csv")
tracking_week_6 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_6.csv")
tracking_week_7 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_7.csv")
tracking_week_8 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_8.csv")
tracking_week_9 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_9.csv")
library(dplyr)

```
Now, I concatenate every week into one dataframe, joined to the tackle dataframe, and clean out unnecessary columns.
```{r}

# Concatenating every week into one dataframe, joined to the tackle dataframe, and cleaning out unnecessary columns

# Assuming 'gameId', 'playId', and 'nflId' are the columns you want to use for the surrogate key
surrogate_key_function <- function(df) {
  df %>%
    mutate(surrogate_key = paste0(gameId, playId, nflId))
}

# Create a list of data frames
totalList <- list(tracking_week_1, tracking_week_2, tracking_week_3, tracking_week_4, tracking_week_5, tracking_week_6, tracking_week_7, tracking_week_8, tracking_week_9)

# Apply the surrogate_key_function to each data frame in totalList
totalList <- lapply(totalList, surrogate_key_function)

# Load the tackles data
tackles <- tackles %>%
  mutate(surrogate_key = paste0(gameId, playId, nflId))

# Initialize an empty dataframe to store the concatenated results
combined_df <- data.frame()

columns_to_remove <- c("forcedFumble", "frameId", "time", "x", "y", "s", "a", "dis", "o", "playDirection", "dir", "jerseyNumber", "pff_missedTackle")

# Loop through each week, perform filtering, and concatenate the results
for (i in seq_along(totalList)) {
  week_t <- inner_join(tackles, totalList[[i]], by = "surrogate_key")
  week_t_filtered <- week_t %>%
    filter(event %in% c("tackle", "out_of_bounds")) %>%
    select(-one_of(columns_to_remove))
  combined_df <- bind_rows(combined_df, week_t_filtered)
}

# Clear the surrogate key and duplicates now that they are no longer needed
combined_df <- combined_df %>%
  select(-c("surrogate_key", "gameId.y", "playId.y", "nflId.y")) %>%
  arrange(gameId.x, nflId.x, playId.x)

# Print the ordered combined dataframe
head(combined_df, 25)


# Calculate missed tackle rate for each defender
defender_missed_tackle_rate <- combined_df |>
  group_by(nflId.x) |>
  summarise(total_tackles = n(),
            total_missed_tackles = sum(event == "tackle"),
            missed_tackle_rate = total_missed_tackles / total_tackles)

# Filter ball carrier tackles where they forced a missed tackle
ball_carrier_forced_missed_tackles <- combined_df |>
  filter(event %in% c("tackle", "out_of_bounds"))

# Join the ball carrier's forced missed tackles with defender missed tackle rates
ball_carrier_defenders_with_high_mtr <- ball_carrier_forced_missed_tackles |>
  group_by(nflId.y) |>
  summarise(defender_missed_tackles = sum(event == "tackle"),
            total_tackles = n()) |>
  left_join(defender_missed_tackle_rate, by = c("nflId.y" = "nflId.x"))

# Print the results
head(ball_carrier_defenders_with_high_mtr, 25)

