---
title: 'Big Data Bowl'
output:
  html_document:
    theme: simplex
    highlight: pygments
    #css: "../css/note-style.css"
  pdf_document: default
---


First, we import all of the data.
```{r}
#getwd()

games <- read.csv("../../nfl-big-data-bowl-2024/games.csv")
players <- read.csv("../../nfl-big-data-bowl-2024/players.csv")
plays <- read.csv("../../nfl-big-data-bowl-2024/plays.csv")
tackles <- read.csv("../../nfl-big-data-bowl-2024/tackles.csv")
tracking_week_1 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_1.csv")
tracking_week_2 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_2.csv")
tracking_week_3 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_3.csv")
tracking_week_4 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_4.csv")
tracking_week_5 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_5.csv")
tracking_week_6 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_6.csv")
tracking_week_7 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_7.csv")
tracking_week_8 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_8.csv")
tracking_week_9 <- read.csv("../../nfl-big-data-bowl-2024/tracking_week_9.csv")
library(dplyr)

```
Now, I concatenate every week into one dataframe, join to the tackle dataframe, and clean out unnecessary columns.
```{r}

library(dplyr)

# Assuming 'gameId', 'playId', and 'nflId' are the columns you want to use for the surrogate key
surrogate_key_function <- function(df) {
  df %>%
    mutate(surrogate_key = paste0(gameId, playId, nflId))
}

# Create a list of data frames
totalList <- list(tracking_week_1, tracking_week_2, tracking_week_3, tracking_week_4, tracking_week_5, tracking_week_6, tracking_week_7, tracking_week_8, tracking_week_9)

# Apply the surrogate_key_function to each data frame in totalList
totalList <- lapply(totalList, surrogate_key_function)

# Load the tackles data
tackles <- tackles %>%
  mutate(surrogate_key = paste0(gameId, playId, nflId))

# Initialize an empty dataframe to store the concatenated results
combined_df <- data.frame()

columns_to_remove <- c("forcedFumble", "frameId", "time", "x", "y", "s", "a", "dis", "o", "playDirection", "dir", "jerseyNumber")

# Loop through each week, perform filtering, and concatenate the results
for (i in seq_along(totalList)) {
  week_t <- inner_join(tackles, totalList[[i]], by = "surrogate_key")
  week_t_filtered <- distinct(week_t)
  week_t_cleaned <- week_t_filtered %>% select(-columns_to_remove)
  combined_df <- bind_rows(combined_df, week_t_cleaned)
}

# Clear the surrogate key and duplicates now that they are no longer needed

col <- c("surrogate_key", "gameId.y", "playId.y", "nflId.y", "event")
combined_df <- combined_df %>% select(-col)
# Arrange the combined dataframe
baseframe <- combined_df %>% arrange(gameId.x, nflId.x, playId.x)
names(baseframe) <- c("gameId", "playId", "nflId", "tackle", "assist", "missedTackle", "displayName", "club")
baseframe <- distinct(baseframe)
# Print the ordered combined dataframe
head(baseframe, 55)


```


Now, I will calculate tackling percentage for each player. First, I'll count their number of attempted tackles (includes successful, assisted, and missed tackles) in the sample set.

```{r}


#names(name_counts) <- c("playerName", "tackleAttempts")
#print(name_counts)


attemptFrame <- baseframe %>%
  count(displayName)





tackleF <- baseframe %>%
  filter(tackle == 1) %>%
  count(displayName)

assistF <- baseframe %>%
  filter(assist == 1) %>%
  count(displayName)

missedF <- baseframe %>%
  filter(missedTackle == 1) %>%
  count(displayName)


resultF <- inner_join(tackleF, assistF, by = "displayName")
resultF <- inner_join(resultF, missedF, by = "displayName")

names(resultF) <- c("displayName", "tackles", "assists", "misses")

resultF <- resultF %>% mutate(attempted = tackles + assists + misses) ### BECAUSE SOME PLAYS A GUY DOES 2


resultF <- resultF %>% arrange(displayName)
head(resultF,25)
```



Now, I'll calculate their tackle percentage. Before doing this, I'll calculate their weighted tackle total, which counts assists as half a tackle. I'll also drop those with <25 tackle attempts.

```{r}
resultF <- resultF %>% mutate(weightedTack = tackles + .5*assists)

resultF <- resultF %>% mutate(tacklePerc = weightedTack / attempted)
resultF <- resultF[resultF$attempted >= 25, , drop = FALSE]
#head(resultF, 25)
resultF
```

Now, I will calculate ball-carriers' weighted forced missed tackle percentage (FMT). I will do this by joining the Plays database to my baseframe, then following similarly to how I calculated Weighted Tackle Percentage.

```{r}

new_base <- baseframe %>% mutate(surrogate_key = paste0(gameId, playId))
new_plays <- plays %>% mutate(surrogate_key = paste0(gameId, playId))

oFrame <- left_join(new_base, new_plays, by = "surrogate_key")

selected <- c("gameId.x", "playId.x", "nflId", "tackle", "assist", "missedTackle", "displayName", "ballCarrierId", "ballCarrierDisplayName", "surrogate_key")

new_base <- oFrame[selected]

#col <- c("gameId.y", "playDescription", "quarter", "down", "yardsToGo", "possessionTeam", "defensiveTeam", "yardlineSide", "yardlineNumber", "gameClock", "preSnapHomeScore", "preSnapVisitorScore", "club", "passResult", "passLength", "prePenaltyPlayResult", "penaltyYards", "playResult", "playNullifiedByPenalty", "absoluteYardlineNumber", "offensiveFormation", "defendersInTheBox")
#oFrame <- oFrame %>% select(-col)

names(new_base) <- c("gameId", "playId", "nflId", "tackle", "assist", "missedTackle", "displayName", "ballCarrierId", "ballCarrierDisplayName", "surrogate_key")
print(new_base)


```

```{r}

selected_column <- "surrogate_key"

# Create a new dataframe with only the selected column
resultOF <- data.frame(surrogate_key = new_base[[selected_column]])


O_attemptFrame <- new_base %>%
  count(ballCarrierDisplayName)

O_tackleF <- new_base %>%
  filter(tackle == 1) %>%
  count(ballCarrierDisplayName)

#names(O_tackleF) <- c("ballCarrierDisplayName", "tackle")
#O_assistF <- new_base %>%
  #filter(assist == 1) %>%
  #count(ballCarrierDisplayName)

#####
O_assistF <- new_base %>%
  filter(assist == 1) %>%
  distinct(ballCarrierDisplayName, surrogate_key) %>%
  group_by(ballCarrierDisplayName) %>%                  ### SO WE DON'T COUNT 5 GUYS TACKLING A CARRIER AS 5 UNIQUE TACKLES
  summarise(assists = n())

#####

O_missedF <- new_base %>%
  filter(missedTackle == 1) %>%
  count(ballCarrierDisplayName)


resultOF <- inner_join(O_tackleF, O_assistF, by = "ballCarrierDisplayName")
resultOF <- inner_join(resultOF, O_missedF, by = "ballCarrierDisplayName")

names(resultOF) <- c("ballCarrierDisplayName", "tackles", "assisted_tackles", "misses")

resultOF <- resultOF %>% mutate(attempted = tackles + assisted_tackles + misses) ### BECAUSE SOME PLAYS A GUY DOES 2


resultOF <- resultOF %>% arrange(ballCarrierDisplayName)
head(resultOF,25)
```
```{r}
new_resultOF <- resultOF

new_resultOF <- new_resultOF %>% mutate(FMT = misses / attempted)

col <- c("tackles", "assisted_tackles")
new_resultOF <- new_resultOF %>% select(-col)

head(new_resultOF, 25)
```

Now, I can calculate tackling efficiency by measuring how defenders tackle hard-to-tackle ball carriers. This will be done by weighting every tackling by the ball carrier's FMT. We will first rule out anyone below a minimum of 10 tackles attempted on them within the sample set. This will help ensure there are no outliers who have exceedingly low or high FMTs. I'll use this table:

```{r}
controlled_FMT <- subset(new_resultOF, attempted >= 10)  

head(controlled_FMT, 25)
```


```{r}
tack_eff <- left_join(new_base, controlled_FMT, by = "ballCarrierDisplayName") 

col <- c("misses", "attempted", "surrogate_key", "ballCarrierId")     #might want to keep bcid in
tack_eff <- tack_eff %>% select(-col)

tack_eff <- tack_eff %>% arrange(displayName)
head(tack_eff, 200)
```
Now, I will weight each tackle by the ball-carrier's FMT. This will allow us to calculate tacklers' percentile scores.


```{r}
custom_count <- function(data, group_column, value_column) {
  result <- data %>%
    group_by({{group_column}}) %>%
    filter(tackle==1) %>% 
    summarise(Tsum = sum({{value_column}}, na.rm = TRUE), .groups = "drop")
  return(result)
}

custom_count_assist <- function(data, group_column, value_column) {
  result2 <- data %>%
    group_by({{group_column}}) %>%
    summarise(Asum = ifelse(any(assist == 1), sum({{value_column}}, na.rm = TRUE), 0), .groups = "drop")
  return(result2)
}


result <- custom_count(tack_eff, displayName, FMT)
result2 <- custom_count_assist(tack_eff, displayName, FMT)
result2 <- result2 %>%
  mutate_all(~ifelse(is.na(.), 0, .))
result3 <- left_join(result, result2, by = "displayName")
result3 <- result3 %>% mutate (sum = Tsum + .5* Asum)
print(result3)

```

```{r}


final_result <- left_join(resultF, result3, by = "displayName")

final_result <- final_result %>% mutate(DATE = sum / attempted)
  final_result <- final_result %>% mutate(DIFF = DATE - tacklePerc)


#final_result <- final_result[final_result$attempted >= 25, , drop = FALSE]
final_result <- final_result %>% arrange(desc(DATE))
col <- c("tackles", "assists", "misses", "attempted", "Tsum", "Asum")
final_result <- final_result %>% select(-col)
#final_result <- final_result %>% arrange(displayName)
print(final_result)
#head(final_result, 50)
```

Now, I can calculate OATE by weighting each Forced Missed Tackle by the Weighted Tackle percentage of the tackler.
```{r}


OATE_base <- left_join(new_base, resultF, by = "displayName")

col <- c("tackles", "assists", "misses", "attempted", "nflId", "playId", "gameId")

OATE_base <- OATE_base %>% select(-col)

controlled_OATE <- subset(OATE_base, weightedTack > 0)

# Reset the row indices
controlled_OATE <- controlled_OATE[, , drop = FALSE]
rownames(controlled_OATE) <- 1:nrow(controlled_OATE)

# Print the result
print(controlled_OATE)

```

```{r}
custom_count <- function(data, group_column, value_column) {
  result <- data %>%
    group_by({{group_column}}) %>%
    filter(missedTackle==1) %>% 
    summarise(sum = sum({{value_column}}, na.rm = TRUE), .groups = "drop")
  return(result)
}


result <- custom_count(controlled_OATE, ballCarrierDisplayName, tacklePerc)

print(result)
```


```{r}
O_final_result <- left_join(new_resultOF, result, by = "ballCarrierDisplayName")

O_final_result <- O_final_result %>% mutate(OATE = sum / attempted)
O_final_result <- O_final_result %>% mutate(DIFF = OATE - FMT)

#final_result <- final_result[final_result$attempted >= 25, , drop = FALSE]
O_final_result <- O_final_result %>% arrange(desc(OATE))
col <- c("Tsum", "Asum")
#O_final_result <- O_final_result %>% select(-col)
#final_result <- final_result %>% arrange(displayName)

cfr2 <- subset(O_final_result, sum > 0)
cfr <- subset(cfr2, misses > 3)

cfr <- cfr[, , drop = FALSE]
rownames(cfr) <- 1:nrow(cfr)

print(cfr)
#head(final_result, 50)
```

```{r}
library(magrittr)
library(ggplot2)
library(ggrepel)
```

```{r}

# Fraction of data to label (e.g., 25%)
library(ggplot2)
library(ggrepel)
library(dplyr)

# Fraction of data to label (e.g., 5%)
label_fraction <- 0.05

# Create a subset of the data with approximately 10% of the rows
label_subset <- cfr %>% sample_frac(label_fraction)

# Identify rows with maximum and minimum values
max_row <- cfr[which.max(cfr$OATE), ]
min_row <- cfr[which.min(cfr$OATE), ]

scatter_plot <- cfr %>%
  ggplot(aes(x = FMT, y = OATE)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_text_repel(data = label_subset,
                  aes(label = ballCarrierDisplayName),
                  size = 3,
                  box.padding = 1,  # Adjust the box padding
                  point.padding = 1) +  # Adjust the point padding
  geom_text_repel(data = rbind(max_row, min_row),
                  aes(label = ballCarrierDisplayName),
                  size = 3,
                  box.padding = 1,  # Adjust the box padding
                  point.padding = 1,  # Adjust the point padding
                  color = "red") +
  labs(x = "FMT", y = "OATE") +
  geom_vline(xintercept = mean(cfr$FMT, na.rm = TRUE), linetype = "dashed", color = "red") +
  geom_hline(yintercept = mean(cfr$OATE, na.rm = TRUE), linetype = "dashed", color = "red") +
  theme_minimal() +
  coord_cartesian(xlim = c(min(cfr$FMT, na.rm = TRUE), max(cfr$FMT, na.rm = TRUE)),
                  ylim = c(min(cfr$OATE, na.rm = TRUE), max(cfr$OATE, na.rm = TRUE)))

# Print the scatter plot
print(scatter_plot)

```

```{r}
# Fraction of data to label (e.g., 5%)
label_fraction <- 0.02

# Create a subset of the data with approximately 5% of the rows
label_subset <- final_result %>% sample_frac(label_fraction)

# Identify rows with maximum and minimum values
max_row <- final_result[which.max(final_result$DATE), ]
min_row <- final_result[which.min(final_result$DATE), ]

scatter_plot2 <- final_result %>%
  ggplot(aes(x = tacklePerc, y = DATE)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_text_repel(data = label_subset,
                  aes(label = displayName),
                  size = 3,
                  box.padding = 1,  # Adjust the box padding
                  point.padding = 1) +  # Adjust the point padding
  geom_text_repel(data = rbind(max_row, min_row),
                  aes(label = displayName),  # Corrected the typo here
                  size = 3,
                  box.padding = 1,  # Adjust the box padding
                  point.padding = 1,  # Adjust the point padding
                  color = "red") +
  labs(x = "Tackle %", y = "DATE") +
  geom_vline(xintercept = mean(final_result$tacklePerc, na.rm = TRUE), linetype = "dashed", color = "red") +
  geom_hline(yintercept = mean(final_result$DATE, na.rm = TRUE), linetype = "dashed", color = "red") +
  theme_minimal() +
  coord_cartesian(xlim = c(min(final_result$tacklePerc, na.rm = TRUE), max(final_result$tacklePerc, na.rm = TRUE)),
                  ylim = c(min(final_result$DATE, na.rm = TRUE), max(final_result$DATE, na.rm = TRUE)))

# Print the scatter plot
print(scatter_plot2)

```


